Secure Composition of Robust and Optimising Compilers

### 2024-10-10 tech/main.tex
tech

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[a4paper,12pt]{article}
\usepackage[left=1.5cm,top=1.5cm,right=1.5cm,bottom=1.5cm]{geometry}
\usepackage{parskip}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
\section{Case Study}

>>>>

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[a4paper,12pt]{article}
\usepackage[left=1.0cm,top=1.5cm,right=1.0cm,bottom=1.5cm]{geometry}
\usepackage{parskip}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
\section{Case Study}\label{sec-cs}
This part instantiates the previously presented theory. 
\Cref{subsec-cs-properties} presents all properties.

\subsection{Properties}\label{subsec-cs-properties}

The properties considered in this case-study are a mixture of trace-properties as welll as hyperproperties.
However, to carry out proofs, all properties are overapproximated by a monitor.

\subsubsection{Formal Definition}\label{subsubsec-cs-properties-formaldef}

\paragraph{Temporal Memory Safety}

\gls*{tms}~\cite{nagarakatte2010cets} is a safety property that describes that an unallocated object must not be (re-)used.

\begin{definition}{\definitionlabel[\glsfirst{tms}]{def-trace-tmsdef}}
  \[
  \tmssafe:=\left\{\varTrace[_{\mssafe}]
    \left| 
      \begin{array}{rcl}
        \ev{Alloc\ \varLoc\ n}&\le_{\varTrace[_{\mssafe}]}&\ev{Dealloc\ \varLoc} \\
        \ev{Use\ \varLoc\ n}&\le_{\varTrace[_{\mssafe}]}&\ev{Dealloc\ \varLoc} \\
        \text{at most one }\ev{Dealloc\ \varLoc}&\text{in}&\varTrace[_{\mssafe}] \\
        \text{at most one }\ev{Alloc\ \varLoc\ n}&\text{in}&\varTrace[_{\mssafe}] \\
      \end{array}
    \right.
  \right\}
  \]
\end{definition}
Hereby, the notation $\varEvent[_{1}]\le_{\varTrace}\varEvent[_{2}]$ means that if $\varEvent[_{1}]$ is in $\varTrace$ and if $\varEvent[_{2}]$ is in $\varTrace$, then $\varEvent[_{1}]$ appears before $\varEvent[_{2}]$.

\paragraph{Spatial Memory Safety}
\gls*{sms}~\cite{nagarakatte2009soft} is a safety property that prohibits out-of-bounds accesses.

\begin{definition}{\definitionlabel[\glsfirst{sms}]{def-trace-smsdef}}
  \[
  \smssafe:=\left\{\varTrace[_{\mssafe}] \left|\begin{array}{rcl}
      \text{If }\ev{Alloc\ \varLoc\ n}\le_{\varTrace_{\mssafe}}\ev{Use\ \varLoc\ m}, \text{ then }m<n
  \end{array}\right.\right\}
  \]
\end{definition}

\paragraph{Memory Safety}

In spirit of earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}, full \gls*{ms} is the intersection of \Cref{def:trace:tmsdef,def:trace:smsdef}.

\begin{definition}{\definitionlabel[\glsfirst{ms}]{def-trace-msdef}}
  \[
  \mssafe:=\tmssafe \cap \smssafe
  \]
\end{definition}

Note that \Cref{def:trace:msdef} ignores data isolation, so there may still be memory-safety issues introduced by side-channels.

\subsubsection{Strict Cryptographic Constant Time}

\gls*{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
This is because, intuitively, hypersafety properties can relate multiple execution traces with each other, but monitors work on a single execution.
It is a common trick to sidestep this issue by means of overapproximation: this section defines the property \gls*{scct}, a stricter variant of \gls*{cct} (inspired by earlier work~\cite{almeida2017jasmin}) that enforces the policy that no secret appears on a trace.
Programs that satisfy \gls*{scct} also satisfy \gls*{cct}, but programs that satisfy \gls*{cct} may not satisfy \gls*{scct}.

\begin{definition}{\definitionlabel[\glsfirst{scct}]{def-trace-scctdef}}
  \noindent\[
  \scctsafe:=\left\{\varTrace[_{\ctsafe}] 
      \left|
        \begin{array}{l}
          \varTrace[_{\ctsafe}]=\hole{\cdot} \text{ or } \\
          \exists\varTrace[_{\ctsafe}'],\varTrace[_{\ctsafe}]=\varEvent[_{\ctsafe}];\unlock\cdot\varTrace[_{\ctsafe}'] \wedge \varTrace[_{\ctsafe}']\in\scctsafe
        \end{array}
      \right.
    \right\}
  \]
\end{definition}

\gls*{scct} may appear overly strict, since it seems that secrets must not occur on a trace (since $\varSecuritytag$ is forced to be $\unlock$). 
However, this is considered standard practice in terms of coding guidelines~\cite{ctguidelines}.
Moreover, programs that have been compiled with FaCT~\cite{cauligi2019fact} and run with a ``data independent timing mode''~\cite{arm-refman,intel-refman} enabled do not leak secrets.

\paragraph{\gls*{ms}, Strict Cryptographic Constant Time}

The combination of \gls*{ms} and \gls*{scct} is the intersection of these properties, \gls*{msscct}.
However, \gls*{ms} uses a different trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
To remedy this issue, we introduce $\sim_{\ctsafe}: \preevent[_{\ctsafe}] \times \preevent[_{\mssafe}] $, a cross-language trace relation (whose key cases are presented below), that we use to intuitively unify the trace model in which the two properties are expressed:
\[
  \typerulenolabel{}{\varEvent[_{\mssafe}]\sim_{\ctsafe}\varEvent[_{\mssafe}];\unlock{}}
  \hspace{2ex}
  \typerulenolabel{}{\lightning\sim_{\ctsafe}\lightning}
  \hspace{2ex}
  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Branch\;n};\varSecuritytag{}}
  \hspace{2ex}
  \typerulenolabel{}{\emptyevent\sim_{\ctsafe}\ev{Binop\;n};\varSecuritytag{}}
\]

Essentially, $\sim_{\ctsafe}$ ignores both the new $\ev{Branch\ n}$ and $\ev{Binop\ n}$ base-events as it relates security-insensitive actions ($\unlock$) to their equivalent counterparts.
Thus, \gls*{ms} traces trivially satisfy \gls*{scct}.
% 
The above relation is extended point-wise to traces, skipping the empty event $\emptyevent$ on either side, and it is now possible to define \gls*{msscct} using the universal image:

\begin{definition}{\definitionlabel[\glsfirst{ms} and \glsfirst{scct}]{def-trace-msscctdef}}
  \[
  \msscctsafe:=\mssafe\cap\sigma_{\sim_{\ctsafe}}\left(\scctsafe\right)
  \]
\end{definition}

\paragraph{Extending the Trace Model with Speculation}

So far, the considered trace models do not let us express speculative execution attacks such as Spectre~\cite{kocher2019spectre}. 
For this, we extend the earlier trace model so that the security tags ($\varSecuritytag{}$) carry additional information about the kind of private data leakage, i.e., the type of speculative leak.
Moreover, we add base-events signalling the beginning of a speculative execution ($\ev{Spec}$), a barrier ($\ev{Barrier}$) that signals that any speculative execution may not go past it, as well as a rollback event ($\ev{Rlb}$), which signals that execution resumes to where speculation started.

\vspace{-1em}
{
\[
  \begin{array}{rrcl}
    (\text{Base-Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\ctsafe}] \\
    (\text{Spectre Variants}) & vX &:=& \operatorname{NONE} \mid \operatorname{PHT} \\
    (\text{Security Tags}) & \varSecuritytag{} &:=& \lock_{vX} \mid \unlock\\ 
    (\text{Events}) & \varEvent[_{\mathghost}] &:=& \varEvent[_{\mathghost}];\varSecuritytag{} \mid \emptyevent \mid \lightning \mid \ev{Spec} \mid \ev{Rlb} \mid \ev{Barrier}\\ 
  \end{array}
\]
}

Even though the considered Spectre variants are just SPECTRE-PHT~\cite{kocher2019spectre}, NONE just describes secret data as in \gls*{scct} (see \Cref{subsec:scct:tracemodel}), the trace model is general enough to allow for potential future extension with different variants~\cite{kocher2019spectre,maisuradze2018ret2spec,horn2019zero}.
%For sake of readability, this paper just uses the notation $\lock$ in place of $\lock_{\text{PHT}}$.

\subsubsection{Speculative Safety}

\gls*{ss}~\cite{patrignani2021exorcising}, similar to \gls*{scct}, is a sound overapproximation of a variant of noninterference.

\begin{definition}{\definitionlabel[\glsfirst{ss}]{def-trace-ssdef}}
  \noindent

  \begin{nscenter}
    \[
      \sssafe := \left\{
        \varTrace[_{\mathghost}] 
          \left|
            \begin{array}{l}
              \varTrace[_{\mathghost}]=\hole{\cdot} 
                \text{ or } 
              \exists\varTrace[_{\mathghost}'].\\
              %
              \left(\varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\unlock\cdot\varTrace[_{\mathghost}'] 
                \text{or }
              \varTrace[_{\mathghost}]=\varEvent[_{\mathghost}];\lock_{\text{NONE}}\cdot\varTrace[_{\mathghost}']\right)\\
              %
              \text{and }%
              \ \varTrace[_{\mathghost}']\in\sssafe
            \end{array}
          \right.
        \right\}
    \] 
  \end{nscenter}
\end{definition}
The technical setup so far leads to the above definition, where only locks annotated with $\text{SPECTRE-PHT}$ are disallowed to occur on the trace.
That way, programs attaining \gls*{ss} do not necessarily attain \gls*{scct}.

\subsubsection{Speculation Memory Safety}\label{sec:spec-ms-rel}

As before, we need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
To this end, let $\sim_{\mathghost}: \varEvent[_{\mathghost}] \times \varEvent[_{\ctsafe}]$ be a cross-language trace relation whose key cases are below.
The intuition is that \gls*{ss} is trivially satisfied in \gls*{scct}, since speculation is inexpressible there, which amounts to dropping events $\ev{Spec}$, $\ev{Rlb}$, or $\ev{Barrier}$, as well as all base events tagged with $\lock_{\text{PHT}}$.

\[
  \typerulenolabel{}{\varEvent[_{\ctsafe}];\lock\sim_{\mathghost}\varEvent[_{\ctsafe}];\lock_{\text{NONE}}}
  \hspace{2ex}
  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\varEvent[_{\mathghost}];\lock_{\text{PHT}}}
  \hspace{2ex}
  \typerulenolabel{}{\varEvent[_{\ctsafe}];\unlock\sim_{\mathghost}\varEvent[_{\ctsafe}];\unlock}
  \hspace{2ex}
  \typerulenolabel{}{\lightning\sim_{\mathghost}\lightning}
  \hspace{2ex}
  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
  \hspace{2ex}
  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
  \hspace{2ex}
  \typerulenolabel{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
\]

We conclude by defining the ultimate property of interest for secure compilers: \gls*{specms}.
\begin{definition}{\definitionlabel[\glsfirst{specms}]{def-trace-specmsdef}}
  \[
  \specmssafe := \msscctsafe\cap\sigma_{\sim_{\mathghost}}\left(\sssafe\right)
  \]
\end{definition}

\subsubsection{Monitors}\label{subsubsec-cs-properties-monitors}
